---
title: "larvae_settlarvae <- read.csv("lement_mhw"
author: "JT Miller"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in packages
```{r}
library(tidyverse)
library(heatwaveR)
library(ggplot2)
```

```{r}
larvae <- read.csv("Invertebrate_Settlement_All_Years_20211004.csv", header=T, sep = ",", na.strings="-99999")
```

# Lets clean up the data
```{r}
larvae %>% 
  dplyr::select(SITE, DATE_RETRIEVED, DURATION, ARTHROPODA, BIVALVIA, GASTROPODA, OPHIUROIDEA_ASTERPODEA, S_PURPURATUS, M_FRANCISCANUS, TOTAL_URCHINS)
```
### Lets look at each group of larvae and divide out by duration. 
```{r}
larvae_p_day <- larvae %>%
  mutate(arthropods_per_day = (ARTHROPODA/DURATION)) %>% 
  mutate(bivalves_per_day = (BIVALVIA/DURATION)) %>% 
  mutate(gastropods_per_day = (GASTROPODA/DURATION)) %>% 
  mutate(stars_per_day = (OPHIUROIDEA_ASTERPODEA/DURATION)) %>% 
  mutate(purple_urchins_per_day = (S_PURPURATUS/DURATION)) %>% 
  mutate(mussels_per_day = (M_FRANCISCANUS/DURATION)) %>% 
  mutate(total_urchins_per_day = (TOTAL_URCHINS/DURATION))
```

### Now lets take group totals by day 
```{r}
larvae_avg_p_day <- larvae_p_day %>% 
  dplyr::group_by(DATE_RETRIEVED) %>% 
  mutate(avg_arthropods = mean(arthropods_per_day)) %>% 
  mutate(avg_bivalves = mean(bivalves_per_day)) %>% 
  mutate(avg_gastropods = mean(gastropods_per_day)) %>% 
  mutate(avg_stars = mean(stars_per_day)) %>% 
  mutate(avg_purps = mean(purple_urchins_per_day)) %>% 
  mutate(avg_mussels = mean(mussels_per_day)) %>% 
  mutate(avg_urchins = mean(total_urchins_per_day))

#And get rid of the extras

larvae_distinct <- larvae_avg_p_day %>% 
  distinct(DATE_RETRIEVED, .keep_all = TRUE)
```

### Lets try graphing one of the groups
```{r}
ggplot(data = larvae_distinct, aes(x = DATE_RETRIEVED, y = avg_arthropods, group = 1)) +
  geom_line()
#Well this looks reasonably interesting, however it really isnt in a readable form. Lets try to condense the measurements 
```
### Lets sum everything up by months and plot by month-year
```{r}
larvae_distinct$DATE_RETRIEVED_NEW <- strptime(as.character(larvae_distinct$DATE_RETRIEVED), "%m/%d/%Y")

format(larvae_distinct$DATE_RETRIEVED_NEW, "%m-%d-%Y")

larvae_distinct_df <- larvae_distinct %>% 
  mutate(month = format(DATE_RETRIEVED_NEW, "%m"), year = format(DATE_RETRIEVED_NEW, "%Y"), month_Year = format(DATE_RETRIEVED_NEW, "%Y-%m")) 

larvae_distinct_month <- larvae_distinct_df %>% 
  group_by(month, year) %>% 
  mutate(avg_arthropods_p_month = sum(avg_arthropods)) %>% 
  mutate(avg_bivalves_p_month = sum(avg_bivalves)) %>% 
  mutate(avg_gastropods_p_month = sum(avg_gastropods)) %>% 
  mutate(avg_stars_p_month = sum(avg_stars)) %>% 
  mutate(avg_purps_p_month = sum(avg_purps)) %>% 
  mutate(avg_mussels_p_month = sum(avg_mussels)) %>% 
  mutate(avg_urchins_p_month = sum(avg_urchins))

larvae_simplified <- larvae_distinct_month %>% 
  distinct(month_Year, .keep_all = TRUE)
```

### It seems that all of the data is going to have a heavy left skew due to the nature of high low quantity counts, therefore lets do a log transformation to normalize the data. 
```{r}
larvae_simplified2 <- larvae_simplified %>% 
  mutate(log_avg_arthropods_p_month = log10(avg_arthropods_p_month)) %>% 
  mutate(log_avg_bivalves_p_month = log10(avg_bivalves_p_month)) %>% 
  mutate(log_avg_gastropods_p_month = log10(avg_gastropods_p_month)) %>% 
  mutate(log_avg_stars_p_month = log10(avg_stars_p_month)) %>% 
  mutate(log_avg_purps_p_month = log10(avg_purps_p_month)) %>% 
  mutate(log_avg_mussels_p_month = log10(avg_mussels_p_month)) %>% 
  mutate(log_avg_urchins_p_month = log10(avg_urchins_p_month))

```


```{r}
ggplot(data = larvae_simplified, aes(x = month_Year, y = avg_arthropods_p_month, group = 1)) +
  geom_line()

#Weird how theres spikes in the data and all of a sudden just nothing...

ggplot(larvae_simplified, aes(x = avg_arthropods_p_month)) +
  geom_histogram()

ggplot(data=larvae_simplified2, aes(x = log_avg_arthropods_p_month)) +
  geom_histogram()



#A log transformation makes the data more normal
```
```{r}
#lets try honing in on that huge spike and see whats going on in that region. 
larvae_simplified_spike <- larvae_simplified %>% 
  filter(year %in% 1998:2000)

ggplot(larvae_simplified_spike, aes(x = month_Year, y = avg_arthropods_p_month, group = 1)) +
  geom_line() + 
  theme(axis.text.x = element_text(size = 6, angle = 90, vjust = 0.5,))
```

```{r}
ggplot(data = larvae_simplified, aes(x = month_Year, y = avg_urchins_p_month, group = 1)) +
  geom_line()
#Interesting oscillatory pattern
```
```{r}
larvae_simplified_urchin_spikes <- larvae_simplified %>% 
  filter(year %in% 1998:2000)

ggplot(larvae_simplified_spike, aes(x = month_Year, y = avg_arthropods_p_month, group = 1)) +
  geom_line() + 
  theme(axis.text.x = element_text(size = 6, angle = 90, vjust = 0.5,))
```


