---
title: "SBC_LTER_MHW"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This markdown is to look at using the package heatwaveR on the SBC_LTER to see if I can identify marine heatwaves

#This particular data set is from nearshore locations along the Northern Channel Islands in the Santa Barbara Channel, its been on going since 2000. 

```{r}
library(ggplot2)
library(dplyr)
library(heatwaveR)
library(tidyverse)
```
### Bringing in the dataset and getting to know it
```{r}
nci_temp <- read.csv("Island_temp_hourly_allsites_allyears_20210218.csv")
```

```{r}
head(nci_temp)
unique(nci_temp$site)
#So we have 21 sites total. Probably going to want to split up the data into each site. 
#Notice that the temp is also recorded by the hour for each day, this doesn't work with the heatwavesR package fxn, therefore lets create a new column that takes the average of each day as the total record? 
```
### Wrangling and cleaning the data
```{r}
is.character(nci_temp$datetime_gmt) #Make sure the date is in character string data type 

nci_temp_sep <- nci_temp%>% 
  separate(datetime_gmt, c("date", "time"), " ") #We need to separate out the data so that times are separated from the data, note that open   " " denotes that the date and time are separated by a space in the column datetime_gmt
```

 
```{r}
#Next, it looks like we're going to need to specify a particular depth range: 14m and only keep these recordings. (14m has the long term values according to the metadata)
nci_temp_14m <- nci_temp_sep %>% 
  dplyr::filter(depth_m == 14)
```




```{r}
#Now lets find the mean temperature for every day and create a new column for it. 
nci_temp_14m_avgT <- nci_temp_14m %>% 
  group_by(date, site) %>% 
  mutate(mean_daily_temp_c = mean(temperature_c))

# Now lets do a sanity check 

(12.62 + 12.66258 + 12.71138)/3 #Nice, the code does what we want. 
```
```{r}
# Now that we have an avg for each day, lets get rid of all the duplicate days by using distinct()
nci_temp_14m_avgT_daily <- nci_temp_14m_avgT %>% 
  distinct(date, .keep_all = TRUE)

#And lets remove the time column since that's not really representative of anything at this point 

nci_temp_14m_avgT_daily <- select(nci_temp_14m_avgT_daily, -c(time, temperature_c))
```

```{r}

```

```{r}
#We can now divide by sites if we would like, lets try looking at the ANS site. 
nci_temp_ANS <- nci_temp_14m_avgT_daily %>% 
  dplyr::filter(site == "ANS")

nci_temp_ANS <- nci_temp_ANS %>% 
  rename(t = date, temp = mean_daily_temp_c) #There seems to be issues with R interpreting the event_line() variable as a fxn? Couldn't find a workaround therefore I just renamed the columns which ended up working.

#Changing the date column to the data type.
nci_temp_ANS <- nci_temp_ANS %>% 
  mutate(t = as.Date(t))

head(nci_temp_ANS)
```





```{r}
#Detect events in a time series
ANS_ts <- ts2clm(nci_temp_ANS,  climatologyPeriod = c("2003-04-11", "2020-09-18"))

ANS_mhw <- detect_event(ANS_ts)
head(ANS_mhw)
```

```{r}
ANS_mhw$event %>% 
  dplyr::ungroup() %>% 
  dplyr::select(event_no, duration, date_start, date_peak, intensity_max, intensity_cumulative) %>% 
  dplyr::arrange(-intensity_max) %>% 
  head(5)
```
```{r}
#Visualizing
event_line(ANS_mhw, spread = 180, metric = "intensity_max", 
           start_date = "2003-04-11", end_date = "2020-09-18")
```
```{r}
lolli_plot(ANS_mhw, metric = "intensity_max")
```
### This is alot of steps to filter and whatnot, maybe I could create a fxn to do these steps so I dont have to copy paste all of this work. 
```{r}
SBC_siter <- function(dat, site=NULL, temp=NULL, date=NULL, date_adjust=NULL) {
  if(!is.null(site)) {
    dat <- dat[dat$site %in% site, ]
  }
  if(is.null(temp)) {
    dat <- dat %>% rename(temp = mean_daily_temp_c)
  }
  if(is.null(date)) {
    dat <- dat %>% rename(t = date)
  }
  if(is.null(date_adjust)){
    dat <- dat %>% mutate(t = as.Date(t))
  }

  
}

test <- SBC_siter(nci_temp_14m_avgT_daily, site="ANS") #Works!
```

### Using the new function we made to look at each site
```{r}
ans <- SBC_siter(nci_temp_14m_avgT_daily, site = "ANS")
bay <- SBC_siter(nci_temp_14m_avgT_daily, site = "BAY") #cant be used too little years of record. 
bea <- SBC_siter(nci_temp_14m_avgT_daily, site = "BEA") #Cant be used 
ctn <- SBC_siter(nci_temp_14m_avgT_daily, site = "CTN")
cuy <- SBC_siter(nci_temp_14m_avgT_daily, site = "CUY")
gul <- SBC_siter(nci_temp_14m_avgT_daily, site = "GUL")
haz <- SBC_siter(nci_temp_14m_avgT_daily, site = "HAZ")
mor <- SBC_siter(nci_temp_14m_avgT_daily, site = "MOR")
pel <- SBC_siter(nci_temp_14m_avgT_daily, site = "PEL")
pro <- SBC_siter(nci_temp_14m_avgT_daily, site = "PRO")
scp <- SBC_siter(nci_temp_14m_avgT_daily, site = "SCP")
smg <- SBC_siter(nci_temp_14m_avgT_daily, site = "SMG")
smn <- SBC_siter(nci_temp_14m_avgT_daily, site = "smn")
sms <- SBC_siter(nci_temp_14m_avgT_daily, site = "SMS")
srn <- SBC_siter(nci_temp_14m_avgT_daily, site = "SRN")
srs <- SBC_siter(nci_temp_14m_avgT_daily, site = "SRS")
trl <- SBC_siter(nci_temp_14m_avgT_daily, site = "TRL")
val <- SBC_siter(nci_temp_14m_avgT_daily, site = "VAL")
wil <- SBC_siter(nci_temp_14m_avgT_daily, site = "WIL")
win <- SBC_siter(nci_temp_14m_avgT_daily, site = "WIN")
```

```{r}
#Maybe a detector fxn is now in order...
mhwer <- function(dat, site=NULL, event=NULL, graph=NULL, lolliplot=NULL ) {
  if(is.null(site)){
    dat <- ts2clm(dat, climatologyPeriod = c(min(dat$t), max(dat$t)))
  }
  if(is.null(event)){
    to_graph <- detect_event(dat)
  }
  if(is.null(graph)){
     mhw_plot <- event_line(to_graph, spread = 180, metric = "intensity_max", start_date = min(dat$t), end_date = max(dat$t))
  }
  if(is.null(lolliplot)){
    lol_plot <- lolli_plot(to_graph, metric = "intensity_max")
  }
  list(mhw_plot, lol_plot)

}
mhwer(ans)


```




### Lets get to plotting!
```{r}
#Detect events in a time series
ctn_ts <- ts2clm(ctn,  climatologyPeriod = c("2009-01-15", "2020-09-20")) #Note: the dates are found by using min and max for ctn$t in the console. 


ctn_mhw <- detect_event(ctn_ts)

#Visualizing
event_line(ctn_mhw, spread = 180, metric = "intensity_max", 
           start_date = "2009-01-15", end_date = "2020-09-20")

lolli_plot(ctn_mhw, metric = "intensity_max")

```
```{r}
#Detect events in a time series
cuy_ts <- ts2clm(cuy,  climatologyPeriod = c("2009-01-15", "2020-09-20")) #Note: the dates are found by using min and max for ctn$t in the console. 


ctn_mhw <- detect_event(ctn_ts)

#Visualizing
event_line(ctn_mhw, spread = 180, metric = "intensity_max", 
           start_date = "2009-01-15", end_date = "2020-09-20")

lolli_plot(ctn_mhw, metric = "intensity_max")

```

### Data citation: Washburn, L, L. Kui. 2021. SBC LTER: Ocean: Ocean hourly temperature at nearshore locations along the Northern Channel Islands in the Santa Barbara Channel, ongoing since 2000 ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/1dfe35ec63376ba1f02172ab92f8032e. Accessed 2021-11-04.